---
title: "Customer Churn Data Analysis"
---
The churn.csv data contains information of several churn and non-churn customers. 

Read churn.csv data.
```{r}
churnData <- read.csv("/Users/Shima/R_projects/customer_churn/churn.csv", header = TRUE)
head(churnData)
```
Clean data before proceeding to create training/testing sets.
```{r}
# check data: count number of null rows
sum(!complete.cases(churnData))

# remove null rows 
churnData <- churnData[complete.cases(churnData), ]

# need to reduce categories before we are able to create a model
# multiple versions of No category: No Internet Service, No Phone Service, binary 0, etc. -> must be consistently No instead

# replace No Internet Service category in internet related columns with No - columns 9-15 are internet related
for(i in 9:15) {
    churnData[churnData[,i]=="No internet service", i] <- "No"
    churnData[,i] <- as.factor(churnData[,i])
}

# replace No Internet Service category in MultipleLines with No 
churnData[churnData$MultipleLines=="No phone service", "MultipleLines"] <- "No"
churnData$MultipleLines <- as.factor(churnData$MultipleLines)

# replace binary categories in SeniorCitizen column with Yes or No 
churnData[churnData$SeniorCitizen=="0", "SeniorCitizen"] <- "No"
churnData[churnData$SeniorCitizen=="1", "SeniorCitizen"] <- "Yes"
churnData$SeniorCitizen <- as.factor(churnData$SeniorCitizen)

head(churnData)

# group tenures to decrease number of categories 
min(churnData$tenure)
max(churnData$tenure)
churnData$tenureGroup <- NA
numRows <- dim(churnData)[1]
for(i in 1:numRows) {
    tenure <- churnData$tenure[i] 
    if (tenure >= 0 & tenure <= 12){
        churnData$tenureGroup[i] <- "0-12 Month"
    }else if(tenure > 12 & tenure <= 24){
        churnData$tenureGroup[i] <- "12-24 Month" 
    }else if (tenure > 24 & tenure <= 48){
        churnData$tenureGroup[i] <- "24-48 Month"
    }else if (tenure > 48 & tenure <=60){
         churnData$tenureGroup[i] <- "48-60 Month" 
    }else if (tenure > 60){
        churnData$tenureGroup[i] <- "> 60 Month"
    }
}
churnData$tenureGroup <- as.factor(churnData$tenureGroup)
```

Randomly select 35% of the data and use that as the test set.
```{r}
# create testing / training datasets
num_samples = dim(churnData)[1]
sampling.rate = 0.35
testing <-sample(1:num_samples,sampling.rate*num_samples,replace = TRUE)
testingset <- subset(churnData[testing,])
training <- setdiff(1:num_samples,testing)
trainingset <- subset(churnData[training,])
```
Create a random forest model to predict customer churn and use this model on the test set.
```{r}
library("randomForest")
randomForestModel <- randomForest(Churn ~. -tenure -customerID, data=trainingset, ntree=1000)
plot(randomForestModel)
legend("top", colnames(randomForestModel$err.rate), col=1:3, cex=0.8,fill=1:3)

# predict on testing set
prdictions <- predict(randomForestModel, testingset)
predictionResults <- data.frame(Churn = prdictions)
head(predictionResults)
```
Evaluate the accuracy of the random forest model over the test set.
```{r}
# test is the one we predict 
wrongPredictions = sum(predictionResults$Churn != trainingset$Churn) 

# claculate misclassification rate
sizeOfTestSet = dim(trainingset)[1]
misclassification = wrongPredictions / sizeOfTestSet
print(paste("Missclassification Rate: ", round(misclassification*100,2), "%"))
```

*Confusion Matrix Setup:*
```{r}
library("crossval")
cm = confusionMatrix(testingset$Churn, predictionResults$Churn, negative="No") 
cm
```
For the test set:  
How many predictions were yes?  
```{r}
print(paste("Number of Yes predictions:",sum(predictionResults$Churn == "Yes")))
print(paste("Number of true positives:",cm[2]))
```
How many predictions were yes but wrong?
```{r}
print(paste("Number of false positives:",cm[1]))
```
How many predictions were no?
```{r}
print(paste("Number of No predictions:",sum(predictionResults$Churn == "No")))
print(paste("Number of true negatives:",cm[3]))
```
How many predictions were no but wrong?
```{r}
print(paste("Number of false negatives:",cm[4]))
```
